---
title: "StatsProject"
output:
  html_document: default
  word_document: default
date: "2024-12-03"
---

# Introduction

This analysis aims to understand how the penalty variants affect the expected profit. We explore statistical methods such as ANOVA, t-tests, and chi-square tests, as well as visualizations like bar graphs and stratification by penalty variants.

### Load Required Libraries

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyverse)
library(readr)
library(ggplot2)
library(pwr)
```

### Load Dataset

```{r, echo=TRUE, message=FALSE, warning=FALSE}

order_file<- read_csv("order_file.csv",
                 col_types = cols(
                   order.id = col_character(),
                   driver.id = col_character(),
                   business.type = col_character(),
                   expected.profit = col_double(),
                   order.placed.time = col_character(),
                   delivery.completed.time = col_character(),
                   cancel.dummy = col_double()
                 ))
penalty_file <- read_csv("penalty_file.csv")
view(order_file)
cat("The dataset has", nrow(order_file), "rows.\n")
view(penalty_file)
cat("The dataset has", nrow(penalty_file), "rows.\n")
```

### Merge Dataset

```{r}
data <- merge(order_file, penalty_file, by = "driver.id")
options(scipen = 999)  # To prevent scientific notation
```

### Clean Data

```{r}
# Remove rows with missing values in 'expected.profit'
data <- data[!is.na(data$expected.profit), ]
```

### Mean and Standard Deviation of expected profit by penalty variant

```{r}
####expected mean of the Expected profit and Penalty variant####
summary_stats <- aggregate(expected.profit ~ penalty.variant, data = data, FUN = function(x) c(mean = mean(x), sd = sd(x)))
summary_stats <- do.call(data.frame, summary_stats) 
names(summary_stats)[2:3] <- c("mean", "sd")
print(summary_stats)
```

### Cancellation Rates by penalty variant

```{r}
# Calculate cancellation rates by penalty variant
cancel_rate_by_penalty <- aggregate(cancel.dummy ~ penalty.variant, data = data, FUN = mean)
cancel_rate_by_penalty$cancellation_rate_pct <- cancel_rate_by_penalty$cancel.dummy * 100
print(cancel_rate_by_penalty)

```

### Max and Min values of cancel dummy with expected profit

```{r}
max_min_summary <- aggregate(expected.profit ~ cancel.dummy, data = data, 
                             FUN = function(x) c(Max = max(x, na.rm = TRUE), Min = min(x, na.rm = TRUE)))
max_min_summary <- do.call(data.frame, max_min_summary)
names(max_min_summary)[2:3] <- c("Maximum", "Minimum")
print(max_min_summary)
```

### Request based vs Rider Based Cancellation Rates

```{r}
# Request-Based Cancellation Rate
request_based_rate <- mean(data$cancel.dummy)
print(paste("Request-Based Cancellation Rate:", round(request_based_rate * 100, 2), "%"))

# Rider-Based Cancellation Rate
# Calculate the cancellation rate for each driver (rider) and then average
rider_based_data <- aggregate(cancel.dummy ~ driver.id, data = data, mean) 
rider_based_rate <- mean(rider_based_data$cancel.dummy)
print(paste("Rider-Based Cancellation Rate:", round(rider_based_rate * 100, 2), "%"))
rate_difference <- (rider_based_rate - request_based_rate) * 100
print(paste("Difference Between Rates:", round(rate_difference, 2), "%"))
```

### Plot of Request_Based and Ride-Based Cancellation

```{r}
# Request-Based Cancellation Rate
request_based_rate <- mean(data$cancel.dummy)
# Rider-Based Cancellation Rate
rider_based_data <- aggregate(cancel.dummy ~ driver.id, data = data, mean)
rider_based_rate <- mean(rider_based_data$cancel.dummy)

cancellation_rates <- data.frame(
  Rate_Type = c("Request-Based", "Rider-Based"),
  Cancellation_Rate = c(request_based_rate * 100, rider_based_rate * 100)
)
ggplot(cancellation_rates, aes(x = Rate_Type, y = Cancellation_Rate, fill = Rate_Type)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  geom_text(aes(label = paste0(round(Cancellation_Rate, 2), "%")), vjust = -0.3) +
  labs(
    title = "Cancellation Rates: Request-Based vs Rider-Based",
    x = "Rate Type",
    y = "Cancellation Rate (%)"
  ) +
  theme_minimal()


```

### Completion rate by Penalty Variant

```{r}
# Calculate completion rate (1 - cancel.dummy) for each penalty variant
completion_rate <- data %>%
  group_by(penalty.variant) %>%
  summarize(Completion_Rate = mean(1 - cancel.dummy) * 100)

# Plot
ggplot(completion_rate, aes(x = penalty.variant, y = Completion_Rate, fill = penalty.variant)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  geom_text(aes(label = paste0(round(Completion_Rate, 2), "%")), vjust = -0.3) +
  labs(
    title = "Completion Rate by Penalty Variant",
    x = "Penalty Variant ($)",
    y = "Completion Rate (%)"
  ) +
  theme_minimal()

```

### Cancellation Rate by Each Business Type

```{r}
# Calculate cancellation rates for each business type
business_cancellation <- data %>%
  group_by(business.type) %>%
  summarize(Cancellation_Rate = mean(cancel.dummy) * 100)
ggplot(business_cancellation, aes(x = business.type, y = Cancellation_Rate, fill = business.type)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  geom_text(aes(label = paste0(round(Cancellation_Rate, 2), "%")), vjust = -0.3) +
  labs(
    title = "Cancellation Rates by Business Type",
    x = "Business Type",
    y = "Cancellation Rate (%)"
  ) +
  theme_minimal()

```

```{r}
# Calculate mean and standard deviation of expected profit for each penalty variant
expected_profit <- data %>%
  group_by(penalty.variant) %>%
  summarize(
    Mean_Profit = mean(expected.profit, na.rm = TRUE),
    SD = sd(expected.profit, na.rm = TRUE)
  )

# Plot
ggplot(expected_profit, aes(x = penalty.variant, y = Mean_Profit, fill = penalty.variant)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  geom_errorbar(aes(ymin = Mean_Profit - SD, ymax = Mean_Profit + SD), width = 0.2) +
  geom_text(aes(label = paste0("$", round(Mean_Profit, 2))), vjust = -0.3) +
  labs(
    title = "Expected Profit by Penalty Variant",
    x = "Penalty Variant ($)",
    y = "Mean Profit ($)"
  ) +
  theme_minimal()

```

### Power Analysis for A/B Testing

```{r}

effect_size <- 0.2  # Small effect size
alpha <- 0.05       # Significance level
power <- 0.8        # Desired power
# Calculate required sample size for a two-sample t-test
sample_size <- pwr.t.test(d = effect_size, sig.level = alpha, power = power, type = "two.sample")$n
print(paste("Required Sample Size per Group:", ceiling(sample_size)))
# Check if current sample size is sufficient
current_sample_size <- nrow(data) / length(unique(data$penalty.variant))
if (current_sample_size >= ceiling(sample_size)) {
  print("The current sample size is sufficient for detecting the effect size.")
} else {
  print("The current sample size is insufficient; consider increasing it.")
}
```

### ANOVA for Penalty Variant

```{r}
#ANOVA to check if penalty.variant affects expected.profit
anova_rides <- aov(expected.profit ~ penalty.variant, data = data)
summary(anova_rides)
```

### Contigency table for cancellations by business type

```{r}
# Create a contingency table for cancellations by business type
table_data <- table(data$business.type, data$cancel.dummy)
row_percentages <- prop.table(table_data, 1) * 100
# Combine the contingency table with row percentages
table_with_percentages <- as.data.frame(table_data)
names(table_with_percentages) <- c("Business Type", "Cancel Dummy", "Frequency")
table_with_percentages$Proportion <- as.vector(row_percentages)
print(table_with_percentages)
```

### Multiple Testing (Bonferroni Correction)

```{r}
bonferroni_test <- pairwise.t.test(data$expected.profit, data$penalty.variant, p.adjust.method = "bonferroni")
bonferroni_test 
```

### t-test for Cancellation Status

```{r}
###expected mean of profit stratified by group pentalty variant###
summary_stats <- aggregate(expected.profit ~ cancel.dummy, data = data, FUN = function(x) c(mean = mean(x), sd = sd(x)))
summary_stats <- do.call(data.frame, summary_stats)
names(summary_stats)[2:3] <- c("mean", "sd") 
print(summary_stats)

##p-value##

t_test_result <- t.test(expected.profit ~ cancel.dummy, data = data, var.equal = FALSE)
print(t_test_result)
```

### Chi-Square Test

```{r}
######Chi-square between dummy and variant####

table_data <- table(data$penalty.variant, data$cancel.dummy)
table_data 

# Perform the Chi-Square Test#
chi_square_result <- chisq.test(table_data)
chi_square_result
```

### Stratification by Penalty variant

```{r}
################stratify this by group###################

data_cntrl <- subset(data, penalty.variant == 0)
data_trt1 <- subset(data, penalty.variant == 10)
data_trt2 <- subset(data, penalty.variant == 20)

###data_cntrl###
summary_stats <- aggregate(expected.profit ~ cancel.dummy, data = data_cntrl, FUN = function(x) c(mean = mean(x), sd = sd(x)))
summary_stats <- do.call(data.frame, summary_stats)  # Format as a data frame
names(summary_stats)[2:3] <- c("mean", "sd")  # Rename columns
summary_stats

t_test_result <- t.test(expected.profit ~ cancel.dummy, data = data_cntrl, var.equal = FALSE)
t_test_result

###data_trt1###
summary_stats <- aggregate(expected.profit ~ cancel.dummy, data = data_trt1, FUN = function(x) c(mean = mean(x), sd = sd(x)))
summary_stats <- do.call(data.frame, summary_stats)  # Format as a data frame
names(summary_stats)[2:3] <- c("mean", "sd")  # Rename columns
summary_stats

t_test_result <- t.test(expected.profit ~ cancel.dummy, data = data_trt1, var.equal = FALSE)
t_test_result

###data_trt2###
summary_stats <- aggregate(expected.profit ~ cancel.dummy, data = data_trt2, FUN = function(x) c(mean = mean(x), sd = sd(x)))
summary_stats <- do.call(data.frame, summary_stats)  # Format as a data frame
names(summary_stats)[2:3] <- c("mean", "sd")  # Rename columns
summary_stats

t_test_result <- t.test(expected.profit ~ cancel.dummy, data = data_trt2, var.equal = FALSE)
t_test_result
```

### Further Analysis because of no converging insights : did chi square between business.type and cancel.dummy

```{r}
######business.type and dummy################ for confounding factor if get any difference from business type

# Create a contingency table
table_data <- table(data$business.type, data$cancel.dummy)
print("Contingency Table:")
print(table_data)

# Calculate Row and Column Percentages
row_percentages <- prop.table(table_data, 1) * 100  
col_percentages <- prop.table(table_data, 2) * 100

# Combine the table with row and column percentages for easy viewing
table_with_percentages <- cbind(table_data, Column_Percentage = col_percentages[, 1])
print("Table with Row and Column Percentages:")
table_with_percentages

# Perform the Chi-Square Test
chi_square_result <- chisq.test(table_data)
chi_square_result
```

### Lets perform same thing with business.type and penalty.variant

```{r}
# Create a contingency table
table_data <- table(data$business.type, data$penalty.variant)
print("Contingency Table:")
print(table_data)

# Calculate Row and column Percentages
row_percentages <- prop.table(table_data, 1) * 100  # Row-wise percentages (1 means rows)
col_percentages <- prop.table(table_data, 2) * 100  # Column-wise percentages (2 means columns)

# Combine the table with row and column percentages for easy viewing
table_with_percentages <- cbind(table_data, Column_Percentage = col_percentages[, 1])
print("Table with Row and Column Percentages:")
table_with_percentages

# Perform the Chi-Square Test
chi_square_result <- chisq.test(table_data)
chi_square_result

```

### Linear Model

```{r}
adjusted_model <- lm(expected.profit ~ factor(business.type) + factor(penalty.variant) + factor(cancel.dummy), data = data)
summary(adjusted_model)
```

### Plots

```{r}
table_data <- table(data$penalty.variant, data$cancel.dummy)
print(table_data)
table_data_col_pct <- prop.table(table_data, 2) * 100 
print(table_data_col_pct)
graph_data <- as.data.frame(table_data_col_pct)
names(graph_data) <- c("Penalty_variant", "Cancel_Dummy", "Percentage")
custom_colors <- c("0" = "darkred", "1" = "darkgreen")
ggplot(graph_data, aes(fill = Cancel_Dummy, y = Percentage, x = Penalty_variant)) +
  geom_bar(position = "dodge", stat = "identity") + 
  scale_fill_manual(values = custom_colors) + 
  ggtitle("Cancellation Requests by Penalty Variant (Column Percentages)") + 
  xlab("Penalty Variant ($)") + 
  ylab("Cancellation Percentage") + 
  geom_text(aes(label = paste0(round(Percentage, 1), "%")),  
            position = position_dodge(0.8), vjust = -0.5) +  
  theme_minimal() + 
  theme(legend.position = "bottom")  
```

### Plot cancellation rates by penalty variant(Row percentage)

```{r}
# Calculate cancellation rates by penalty variant
cancel_rate_by_penalty <- aggregate(cancel.dummy ~ penalty.variant, data = data, FUN = function(x) mean(x))
names(cancel_rate_by_penalty) <- c("penalty.variant", "cancellation_rate")
print(cancel_rate_by_penalty)
ggplot(cancel_rate_by_penalty, aes(x = factor(penalty.variant), y = cancellation_rate, fill = factor(penalty.variant))) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  labs(title = "Cancellation Rate by Penalty Variant(Row Percentage)", x = "Penalty Variant ($)", y = "Cancellation Rate") +
  theme_minimal() +
  geom_text(aes(label = scales::percent(cancellation_rate, accuracy = 0.01)), vjust = -0.3, size = 4)
```

### Profit in overall \$ for each variant

```{r}
# Calculate total profit for each penalty variant
total_profit <- data %>%
  group_by(penalty.variant) %>%
  summarize(Total_Profit = sum(expected.profit, na.rm = TRUE))

# Plot
ggplot(total_profit, aes(x = penalty.variant, y = Total_Profit, fill = penalty.variant)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  geom_text(aes(label = paste0("$", round(Total_Profit, 2))), vjust = -0.3) +
  labs(
    title = "Total Profit by Penalty Variant",
    x = "Penalty Variant ($)",
    y = "Total Profit ($)"
  ) +
  theme_minimal()

```
